Molecular dynamics simulation of synchronization in driven particles

### 2021-03-27 manuscript/manuscript_ajp_version2.tex
updated figs

We confine the particles using a position dependent 
potential energy, called a landscape or substrate.
The landscape is modulated in the $y$-direction
by the periodic function 
 \begin{equation}
   U(y) = U_0 \cos{(2 \pi y / \lambda)},
     \label{eq:ysubstrate}
\end{equation}
where $\lambda=L/N_p$, with $N_p$ equal to the number of periods,
and $U_0$ is a parameter
 that sets the depth of the minima
 with simulation units of energy $E_0$. 
 We plot this function in 
 Fig.~\ref{fig:1_landscape}
 for $N_p = 3$.  In Fig.~\ref{fig:1_landscape}(a) we show 
 the $x$-$y$ plane with a contour plot of $U(y)$ 
 to illustrate
 the 2D potential energy landscape;
 the maxima are colored red and the minima colored blue.
 %[xx make sure that all your figures make sense when printed in b/w xx]

The confining force on the particle %[DM removed subscripts] $i$
 is given by
 \begin{equation}
 \vec{F}^{\ell}(\vec{r}) = - \vec \nabla U(\vec{r}). %DM removed ^{\ell}
 \label{eq:dudr}
 \end{equation}
 %[xx what does the superscript $\ell$ mean? xx]
 where $\ell$ denotes the landscape.
 In Fig.~\ref{fig:1_landscape}(b) we plot  
 $U(y)$ to illustrate how the magnitude
 $|\vec{F}^{\ell}|$ is calculated from the particle position $y$.
 %[My collaborators and I consistently use superscripts for these equations to separate the force superscripts from the particle index subscripts.  This is consistent with the textbook I teach from.  xx suggest that you use subscripts instead of superscripts. We can easily make that change in the next iteration of your manuscript. However, you would need to change the figures, e.g., $\vec{F}_{\rm d}$ $F_{\rm dc}$ $F_{\rm ac}$. One of us didn't like superscripts where he was an undergraduate!  xx]

Particles are subject to an external time-dependent driving force
$\vec{F}^{\rm d}(t)$
applied parallel to the $y$-direction.
We model this force as
\begin{equation}
  \vec{F}^{\rm d}(t) = [F^{\rm dc} + F^{\rm ac} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with 
a constant component $F^{\rm dc}$
and a time dependent component with amplitude $F^{\rm ac}$
with angular frequency $\omega = 2 \pi f$.

The inertia of 
small particles is reduced by interactions
with the fluid particles.\cite{Purcell1977}
We assume 
colloids are
overdamped
so the particles do not accelerate,
that is, they are suspended in a continuous viscous fluid
that dissipates energy supplied externally. 
Newton's second law for an individual particle
is simplified
by the assumption that $\vec{a}$ is zero. 
The overdamped equation of motion for
the velocity $\vec{v}$ of 
an isolated particle is
\begin{equation}
  \eta \vec{v} = \vec{F}^{\ell}(\vec{r}) + \vec{F}^{\rm d}(t).
    \label{eq:motion}
\end{equation}
with friction coefficient $\eta = 1$ in units of $v_0 / F_0$.
%[xx added $v_0$ to table, please check xx] 
The term $-\eta \vec{v}$
is a drag force, which models
energy dissipation due to the fluid. 
We discuss  models for
spheres moving through fluids in 
Problem~\ref{ex:reynolds}.

The equation of motion provides a direct calculation of the velocity
of an individual particle at position $\vec{r}$ 
and simulation time $t$.
The   simulation is controlled by a  loop
which runs to a maximum integer time step.
At each time step
we evaluate the net force on each particle as a function of its position
$\vec{r}(t)$
and then integrate
the equation of motion to move particles
to an updated position.
Because the acceleration is zero,
the integration of the equation of motion
is performed via 
the simple first-order Euler method 
\begin{equation}
  \vec{r}(t+\Delta t) = \vec{v}(t) \Delta t + \vec{r}(t)
    \label{eq:euler}
\end{equation}
for a time step $\Delta t = 0.1\,\tau$.
In 
Problem~\ref{ex:euler}
we describe 
the numerical methods for 
solving differential equations.
%------------------------------------------------------------------
\begin{figure} % [h!]
\centering
\includegraphics[width=\columnwidth]{fig1_landscape.pdf}
\caption{
Schematic of the simulation of a single particle
  driven across a washboard potential energy landscape.
  (a) View of the $x$-$y$ plane. 
  The time-dependent applied driving force $\vec{F}^{\rm d}$
  is parallel to the $y$-axis.
  The landscape is shown with 
  maxima in the potential energy marked in red
  and minima marked in blue.
  A particle is 
  subject to competing forces of the landscape and applied driving force.
  (b) The potential energy function
  along the $y$-axis $U(y)$.  %DM removed ^{\ell}
  The particle in (a) is shown at the same $y$-position.
  The slope of $U(y)$ %[DM removed ^{\ell} use the subscripts with the potential, but will move to superscript xx used subscript here xx]
  is the 
  magnitude of force $\vec{F}^{\ell}$. 
  }
\label{fig:1_landscape}
\end{figure}
%
\section{Mode-locking of a single particle}
% In subsection headings, only the first word is capitalized.
\label{sec:results}
Here we drive 
a single particle 
across the landscape. 
The numerical implementation of the landscape 
is calculated with Eqs.~\eqref{eq:ysubstrate} and \eqref{eq:dudr}  %[DM - right, two negative signs, one from the derivative, the other from the definition of force.  I found the ``mistaken'' dropped minus sign in the depths of january and added it.  xx but you should get a + sign not a minus sign xx] 
\begin{equation}
  \label{eq:force}
  F^\ell_y(y) = A_{p} \sin{(2 \pi y / \lambda)},
\end{equation}
where the force is scaled by the parameter $A_{p} = 2\pi U_0/\lambda$.
In this section we fix the landscape parameters
to $A_{p} = 0.1\,F_0$ 
with $N_p=20$ minima, 
corresponding to a spatial period $\lambda = 2.3\,a_0$.
The competition between the driving force and the landscape potential
can produce a variety of hopping patterns in the particle motion. 
The relative values of $F^{\rm ac}$, $F^{\rm dc}$ and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
If $F^{\rm d}(t) > A_p$, a particle can 
overcome the barrier height of the landscape,
and 
the particle hops between minima in the energy landscape.
If the driving frequency is low,
as in Fig.~\ref{fig:2_Fd_vy_time},  
the driven particle 
moves 
in a pattern 
with the same frequency 
as the time-dependent force $F^{\rm d}(t)$,
but is modulated by the landscape period.
We explore changes in the frequency and different values of $F^{\rm ac}$ 
in Problem~\ref{ex:parameters}.
Here we vary $F^{\rm dc}$ 
while holding the remaining parameters fixed.

@@

In Fig.~\ref{fig:2_Fd_vy_time}(a)
we plot $F^{\rm d}(t)$ 
as a function of time with the
constants $F^{\rm dc}=0.07$, 
$F^{\rm ac}=0.07$ and $f=0.01$ cycles per time unit $\tau$.
The temporal period of the driving force is
$T = 1/f = 100\,\tau$.
In Fig.~\ref{fig:2_Fd_vy_time}(b) 
we show the $y$-position of the particle
as a function of time,
where we have
normalized $y$ by $\lambda$.
The initial particle position is $y = 0$. 
The particle moves
in the positive $y$-direction
through a distance $\Delta y = \lambda$  in a time $T$,
so that  
the average velocity 
$\langle {v}_y \rangle = \lambda f$. 
The inset of Fig.~\ref{fig:2_Fd_vy_time}(b)
shows $y$ 
over one period $100\,\tau < t < 200\,\tau$
with 
the contour plot described
in Fig.~\ref{fig:1_landscape}(a).
The motion is synchronized such that the 
driving force is a maximum when the landscape 
force is minimum,
as shown by the coincidence of the
steep slope of 
$y/\lambda$ 
and the maxima in $F^{\rm d}(t)$.
If $F^{\rm d}(t)$ is large,
the particle moves across the substrate minima,
shown in blue with the contour plot.
The slope of $y/\lambda$ equals zero twice
during a driving period,
indicating zero
forward motion 
due to a coincidence of the negative
driving force and motion over the landscape maxima.
%------------------------------------------------------------------------
\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{fig2_Fd_vy_time.pdf}
\caption{(a) The applied driving force $F^{\rm d}(t)$ 
  with parameters $F^{\rm dc} = 0.07$, $F^{\rm ac} = 0.07$ and $f=0.01$.
  (b) 
  The $y$-position of the driven particle
  normalized by the period of the substrate $\lambda$.
  The substrate strength is $A_p=0.1$.
  The inset   shows
  the $y$-position
  through the second period $100\tau<t<200\tau$
 along with the contour plot depicting
  the landscape potential described in Fig.~\ref{fig:1_landscape}(a).
  }
\label{fig:2_Fd_vy_time}
\end{figure}

To explore the possible hopping patterns,
we sweep through a range of $F^{\rm dc}$ for fixed $F^{\rm ac}$ and $A_p$.
In Fig.~\ref{fig:3_sweep_vyFDC} 
we increase $F^{\rm dc}$ in increments of $0.001\,F_0$,
and 
measure the average velocity $\langle v_y \rangle $ 
as a function of $F^{\rm dc}$.
We also perform the sweep for a non-oscillatory drive $F^{\rm ac} = 0$.
With no oscillating component of the driving force,
  the force-velocity relation increases monotonically 
  above the depinning threshold $F^c$ such that
  \begin{equation}
    \langle v_y \rangle \propto (F^{\rm dc}-F^c)^{-\beta}.
  \end{equation}
  %[xx what is $\beta$? xx]
  where the power $\beta$ varies with the
  system type and
  can be used to identify universality class. \cite{Reichhardt2017}
  The critical force $F^c$ is equal to the maximum substrate force
  $A_p$.

The addition of an ac drive leads
  to the formation of modes.
  A mode is a periodic pattern of hops
  with a constant average particle velocity, $\langle {v}_{y} \rangle$
  over a range of driving forces $F^{\rm dc}$.
  In Fig.~\ref{fig:3_sweep_vyFDC}
  we sweep $F^{\rm dc}$
  with $F^{\rm ac} = 0.07$ and $f=0.01$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle
due to the landscape confinement.  
At 
low $F^{\rm dc}$ the average velocity $\langle v_y \rangle$ is zero.
Because 
$A_p$ is large compared to the extrema of $F^{\rm d}(t)$,
the particle oscillates back and forth
in a single minima with no net velocity, an example of 
a 0:0 mode.
At higher $F^{\rm dc}$ the particle velocity 
$\langle v_{y} \rangle$ increases in steps of uniform height,
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.
We observe a mode of $n=1$
for the range $0.05 < F^{\rm dc} < 0.08$,
$n=2$ for $0.08 < F^{\rm dc} < 0.11$,
$n=3$ for $0.12 < F^{\rm dc} < 0.13$,
$n=4$ for $0.14 < F^{\rm dc} < 0.155$, and 
$n=5$ for $0.155 < F^{\rm dc} < 0.16$.
Higher modes are not visible.
The step width is nonlinear
and depends on the strength of $F^{\rm ac}$
%as a Bessel function 
for this landscape potential.\cite{Reichhardt2000,Juniper2017}
%[DM the step width - simply removed xx what quantity depends on a Bessel function? xx]
These steps, known as Shapiro steps, can have a variety of
interesting patterns
such as a devil's staircase related to chaotic dynamics.\cite{Bak1986}

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{fig3_sweep_vyFDC.pdf}
\caption{Average particle velocity  $\langle v_{y} \rangle$
  as a function of $F^{\rm dc}$.
  We let
  $F^{\rm ac}=0.0$ (blue) and 
  $F^{\rm ac}=0.07$ (orange) with $f = 0.01$ 
  as in Fig.~\ref{fig:2_Fd_vy_time}.
  The value of $F^{\rm dc}$ for the first four steps
  %is given,
  correspond with the
  panels (a-d) 
  %indicating a system
  plotted in Fig.~\ref{fig:4_phase}.
  %[DM - clarified?  xx don't understand meaning of indicating a system
  %plotted in Fig.~\ref{fig:4_phase} xx]
}
\label{fig:3_sweep_vyFDC}
\end{figure}

  To study synchronization patterns, 
  it is useful to compare
  mode-locked quantities 
  in a two-dimensional phase plot. 
  For a driven pendulum confined to a single potential well,
  an appropriate
  phase space is the particle velocity $v_y$ versus the position $y$.  
  For a particle driven 
  through multiple identical wells 
  we define phase variables 
  to account for the net increase in the position.
  The phase position is
  \begin{equation}
    \phi(t) = 2\pi [y(t)-\langle v_y \rangle t]/\lambda,
  \end{equation}
  centered about the average particle displacement $\langle v_y \rangle t$
  and normalized by the substrate period $\lambda$.\cite{Juniper2015}
  The phase velocity is
  \begin{equation}
    \dot{\phi}(t) =2\pi [v_y(t)-\langle v_y \rangle] /\lambda.  
  \end{equation}
  For zero landscape force, 
  the phase velocity  is zero when $F^{\rm d}(t) = F^{\rm dc}$.
  %[DM rewrite]
  If the system
  is strictly mode-locked,
  the particle velocity
  recurs at a particular spatial location, and 
  %If the velocity
  %is mode-locked to 
  %a spatial location,
  %[xx don't understand meaning of the velocity
  %is mode-locked to 
  %a spatial location xx]
  a closed loop appears in  
  phase space. 
  A 1:1 mode appears as a circle or oval. 
  Nodes appear 
  for higher modes,
  sometimes forming figure-eights
  or other recognizable patterns.
  A system that is nearly phase locked
  will appear as an almost closed loop.
  Such quasiperiodic systems are
  not fully synchronized
  so the position-velocity relation
  shifts in time as in Fig.~\ref{fig:4_phase}.

  In Fig.~\ref{fig:4_phase}
  we plot $\dot{\phi}(t)$ versus $\phi(t)$
  for increasing 
  $F^{\rm dc}$, 
  with the remaining parameters fixed as in Fig.~\ref{fig:2_Fd_vy_time}.
 %[xx this sentence not helpful, what is plotted is clear: For clarity the plots are normalized by the prefactor $2\pi$.  xx]
  For
  $F^{\rm dc} = 0.04$
  the phase plot   in Fig.~\ref{fig:4_phase}(a) with is an
  % small [xx not clear what small means? can we eliminate this word? xx]
  asymmetric curve.
  A tail appears
  due to the initial transient
  motion of the particle.
  The particle is confined to a single
  substrate minima,
  and has no net velocity.
  The asymmetry is caused by bias induced by $F^{\rm dc}$.
  In Fig.~\ref{fig:4_phase}(b)
  with $F^{\rm dc} = 0.07$
  the phase loop is a symmetric triangular shape,
  indicating a 1:1 match between the
  particle motion and velocity consistent with 
  Fig.~\ref{fig:2_Fd_vy_time}(a).
  As $F^{\rm dc}$ is
  increased,
  nodes form in the phase diagram, which
   occur due to repeated values
  of $\dot{\phi}$ over multiple phase positions.
  In Fig.~\ref{fig:4_phase}(c)
  with $F^{\rm dc} = 0.1$
  two nodes form.
  The particle moves a distance $2\lambda$
  during one time period.
For $F^{\rm dc} = 0.125$ as in Fig.~\ref{fig:4_phase}(d),
  three nodes form as the particle moves across $3\lambda$.

    \begin{figure}[h!]
      \centering
      \includegraphics[width=\columnwidth]{fig4_phase.pdf}
      \caption{
        Phase plot of $\dot{\phi(t)}/(2\pi)$ versus $\phi(t)/(2\pi)$.
        The particle is driven with $F^{\rm dc}$ equal to  (a) $0.04$, (b) $0.07$, (c) $0.1$, and (d) $0.125$.  These values are denoted as black circles
        in Fig.~\ref{fig:3_sweep_vyFDC}. 
      The other parameters
      are $F^{\rm ac} = 0.07$, $f=0.01$, and $A_p = 0.1$
      as in Fig.~\ref{fig:2_Fd_vy_time}.}
      \label{fig:4_phase}
    \end{figure}

@@

@@

@@

\item [(b)] If $m/\gamma$ is small, the
  particle's acceleration can be ignored
  entirely.
Using Newton's second law for
  a small momentum relaxation time, 
  show that a particle confined to a landscape exerting force
  $F^\ell(\vec{r})$ subject to a time dependent drive $F^{\rm d}(t)$
  can be modeled by the equation of motion described in 
  Eq.~\eqref{eq:motion}.
  \end{enumerate}
\end{problem}

\begin{problem}{Integration methods}
  \label{ex:euler}
%[xx you didn't really discuss other integration algorithms, you just mentioned some names That's OK. Instead you might want to make this problem 1 and guide  readers to write their own program. Then mention that a program is available, which you only mention in the abstract xx]

>>>>

We confine the particles using a position dependent 
potential energy, called a landscape or substrate.
The landscape is modulated in the $y$-direction
by the periodic function 
 \begin{equation}
   U(y) = U_0 \cos{(2 \pi y / \lambda)},
     \label{eq:ysubstrate}
\end{equation}
where $\lambda=L/N_p$, with $N_p$ equal to the number of periods,
and $U_0$ is a parameter
 that sets the depth of the minima
 with simulation units of energy $E_0$. 
 We plot this function in 
 Fig.~\ref{fig:1_landscape}
 for $N_p = 3$.  In Fig.~\ref{fig:1_landscape}(a) we show 
 the $x$-$y$ plane with a contour plot of $U(y)$ 
 to illustrate
 the 2D potential energy landscape;
 the maxima are colored white and the minima colored gray.
 %[DM - switch from red/blue to white/gray - xx make sure that all your figures make sense when printed in b/w xx]

The confining force on the particle %[DM removed subscripts] $i$
 is given by
 \begin{equation}
 \vec{F}_{\ell}(\vec{r}) = - \vec \nabla U(\vec{r}). %DM removed ^{\ell}
 \label{eq:dudr}
 \end{equation}
 %[xx what does the superscript $\ell$ mean? xx]
 where $\ell$ denotes the landscape.
 In Fig.~\ref{fig:1_landscape}(b) we plot  
 $U(y)$ to illustrate how the magnitude
 $|\vec{F}_{\ell}|$ is calculated from the particle position $y$.
 %[DM - begrudingly changed - My collaborators and I consistently use superscripts for these equations to separate the force superscripts from the particle index subscripts.  This is consistent with the textbook I teach from.  xx suggest that you use subscripts instead of superscripts. We can easily make that change in the next iteration of your manuscript. However, you would need to change the figures, e.g., $\vec{F}_{\rm d}$ $F_{\rm dc}$ $F_{\rm ac}$. One of us didn't like superscripts where he was an undergraduate!  xx]

Particles are subject to an external time-dependent driving force
$\vec{F}_{\rm d}(t)$
applied parallel to the $y$-direction.
We model this force as
\begin{equation}
  \vec{F}_{\rm d}(t) = [F_{\rm dc} + F_{\rm ac} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with 
a constant component $F_{\rm dc}$
and a time dependent component with amplitude $F_{\rm ac}$
with angular frequency $\omega = 2 \pi f$.

The inertia of 
small particles is reduced by interactions
with the fluid particles.\cite{Purcell1977}
We assume 
colloids are
overdamped
so the particles do not accelerate,
that is, they are suspended in a continuous viscous fluid
that dissipates energy supplied externally. 
Newton's second law for an individual particle
is simplified
by the assumption that $\vec{a}$ is zero. 
The overdamped equation of motion for
the velocity $\vec{v}$ of 
an isolated particle is
\begin{equation}
  \eta \vec{v} = \vec{F}_{\ell}(\vec{r}) + \vec{F}_{\rm d}(t).
    \label{eq:motion}
\end{equation}
with friction coefficient $\eta = 1$ in units of $v_0 / F_0$.
%[xx added $v_0$ to table, please check xx] 
The term $-\eta \vec{v}$
is a drag force, which models
energy dissipation due to the fluid. 
We discuss  models for
spheres moving through fluids in 
Problem~\ref{ex:reynolds}.

The equation of motion provides a direct calculation of the velocity
of an individual particle at position $\vec{r}$ 
and simulation time $t$.
The   simulation is controlled by a  loop
which runs to a maximum integer time step.
At each time step
we evaluate the net force on each particle as a function of its position
$\vec{r}(t)$
and then integrate
the equation of motion to move particles
to an updated position.
Because the acceleration is zero,
the integration of the equation of motion
is performed via 
the simple first-order Euler method 
\begin{equation}
  \vec{r}(t+\Delta t) = \vec{v}(t) \Delta t + \vec{r}(t)
    \label{eq:euler}
\end{equation}
for a time step $\Delta t = 0.1\,\tau$.
In 
Problem~\ref{ex:euler}
we describe 
the numerical methods for 
solving differential equations.
%------------------------------------------------------------------
\begin{figure} % [h!]
\centering
\includegraphics[width=\columnwidth]{fig1_landscape.pdf}
\caption{
Schematic of the simulation of a single particle
  driven across a washboard potential energy landscape.
  (a) View of the $x$-$y$ plane. 
  The time-dependent applied driving force $\vec{F}_{\rm d}$
  is parallel to the $y$-axis.
  The landscape is shown with
  %[DM changed color scheme from red/blue to gray/white for better BW conversion]
  maxima in the potential energy marked in white %#red
  and minima marked in gray %.
  A particle is 
  subject to competing forces of the landscape and applied driving force.
  (b) The potential energy function
  along the $y$-axis $U(y)$.  %DM removed _{\ell}
  The particle in (a) is shown at the same $y$-position.
  The slope of $U(y)$ %[DM removed _{\ell} use the subscripts with the potential, but will move to superscript xx used subscript here xx]
  is the 
  magnitude of force $\vec{F}_{\ell}$. 
  }
\label{fig:1_landscape}
\end{figure}
%
\section{Mode-locking of a single particle}
% In subsection headings, only the first word is capitalized.
\label{sec:results}
Here we drive 
a single particle 
across the landscape. 
The numerical implementation of the landscape 
is calculated with Eqs.~\eqref{eq:ysubstrate} and \eqref{eq:dudr}  %[DM - right, two negative signs, one from the derivative, the other from the definition of force.  I found the ``mistaken'' dropped minus sign in the depths of january and added it.  xx but you should get a + sign not a minus sign xx] 
\begin{equation}
  \label{eq:force}
  {F_{\ell}}_y(y) = A_{p} \sin{(2 \pi y / \lambda)},
\end{equation}
where the force is scaled by the parameter $A_{p} = 2\pi U_0/\lambda$.
In this section we fix the landscape parameters
to $A_{p} = 0.1\,F_0$ 
with $N_p=20$ minima, 
corresponding to a spatial period $\lambda = 2.3\,a_0$.
The competition between the driving force and the landscape potential
can produce a variety of hopping patterns in the particle motion. 
The relative values of $F_{\rm ac}$, $F_{\rm dc}$ and $A_p$
control the rate and distance a  particle moves 
forward and backward in the landscape.
If $F_{\rm d}(t) > A_p$, a particle can 
overcome the barrier height of the landscape,
and 
the particle hops between minima in the energy landscape.
If the driving frequency is low,
as in Fig.~\ref{fig:2_Fd_vy_time},  
the driven particle 
moves 
in a pattern 
with the same frequency 
as the time-dependent force $F_{\rm d}(t)$,
but is modulated by the landscape period.
We explore changes in the frequency and different values of $F_{\rm ac}$ 
in Problem~\ref{ex:parameters}.
Here we vary $F_{\rm dc}$ 
while holding the remaining parameters fixed.

@@

In Fig.~\ref{fig:2_Fd_vy_time}(a)
we plot $F_{\rm d}(t)$ 
as a function of time with the
constants $F_{\rm dc}=0.07$, 
$F_{\rm ac}=0.07$ and $f=0.01$ cycles per time unit $\tau$.
The temporal period of the driving force is
$T = 1/f = 100\,\tau$.
In Fig.~\ref{fig:2_Fd_vy_time}(b) 
we show the $y$-position of the particle
as a function of time,
where we have
normalized $y$ by $\lambda$.
The initial particle position is $y = 0$. 
The particle moves
in the positive $y$-direction
through a distance $\Delta y = \lambda$  in a time $T$,
so that  
the average velocity 
$\langle {v}_y \rangle = \lambda f$. 
The inset of Fig.~\ref{fig:2_Fd_vy_time}(b)
shows $y$ 
over one period $100\,\tau < t < 200\,\tau$
with 
the contour plot described
in Fig.~\ref{fig:1_landscape}(a).
The motion is synchronized such that the 
driving force is a maximum when the landscape 
force is minimum,
as shown by the coincidence of the
steep slope of 
$y/\lambda$ 
and the maxima in $F_{\rm d}(t)$.
If $F_{\rm d}(t)$ is large,
the particle moves across the substrate minima,
%[DM - color change]
shown in gray with the contour plot.
The slope of $y/\lambda$ equals zero twice
during a driving period,
indicating zero
forward motion 
due to a coincidence of the negative
driving force and motion over the landscape maxima.
%------------------------------------------------------------------------
\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{fig2_Fd_vy_time.pdf}
\caption{(a) The applied driving force $F_{\rm d}(t)$ 
  with parameters $F_{\rm dc} = 0.07$, $F_{\rm ac} = 0.07$ and $f=0.01$.
  (b) 
  The $y$-position of the driven particle
  normalized by the period of the substrate $\lambda$.
  The substrate strength is $A_p=0.1$.
  The inset   shows
  the $y$-position
  through the second period $100\tau<t<200\tau$
 along with the contour plot depicting
  the landscape potential described in Fig.~\ref{fig:1_landscape}(a).
  }
\label{fig:2_Fd_vy_time}
\end{figure}

To explore the possible hopping patterns,
we sweep through a range of $F_{\rm dc}$ for fixed $F_{\rm ac}$ and $A_p$.
In Fig.~\ref{fig:3_sweep_vyFDC} 
we increase $F_{\rm dc}$ in increments of $0.001\,F_0$,
and 
measure the average velocity $\langle v_y \rangle $ 
as a function of $F_{\rm dc}$.
We also perform the sweep for a non-oscillatory drive $F_{\rm ac} = 0$.
With no oscillating component of the driving force,
  the force-velocity relation increases monotonically 
  above the depinning threshold $F_c$ such that
  \begin{equation}
    \langle v_y \rangle \propto (F_{\rm dc}-F_c)^{-\beta}.
  \end{equation}
  %[xx what is $\beta$? xx]
  where the power $\beta$ varies with the
  system type and
  can be used to identify universality class. \cite{Reichhardt2017}
  The critical force $F_c$ is equal to the maximum substrate force
  $A_p$.

The addition of an ac drive leads
  to the formation of modes.
  A mode is a periodic pattern of hops
  with a constant average particle velocity, $\langle {v}_{y} \rangle$
  over a range of driving forces $F_{\rm dc}$.
  In Fig.~\ref{fig:3_sweep_vyFDC}
  we sweep $F_{\rm dc}$
  with $F_{\rm ac} = 0.07$ and $f=0.01$.
Each step represents a different pattern of hops
between substrate minima
performed by the particle
due to the landscape confinement.  
At 
low $F_{\rm dc}$ the average velocity $\langle v_y \rangle$ is zero.
Because 
$A_p$ is large compared to the extrema of $F_{\rm d}(t)$,
the particle oscillates back and forth
in a single minima with no net velocity, an example of 
a 0:0 mode.
At higher $F_{\rm dc}$ the particle velocity 
$\langle v_{y} \rangle$ increases in steps of uniform height,
$\langle v_{y} \rangle = n \lambda f$,
where $n$ is an integer.
We observe a mode of $n=1$
for the range $0.05 < F_{\rm dc} < 0.08$,
$n=2$ for $0.08 < F_{\rm dc} < 0.11$,
$n=3$ for $0.12 < F_{\rm dc} < 0.13$,
$n=4$ for $0.14 < F_{\rm dc} < 0.155$, and 
$n=5$ for $0.155 < F_{\rm dc} < 0.16$.
Higher modes are not visible.
The step width is nonlinear
and depends on the strength of $F_{\rm ac}$
%as a Bessel function 
for this landscape potential.\cite{Reichhardt2000,Juniper2017}
%[DM the step width - simply removed xx what quantity depends on a Bessel function? xx]
These steps, known as Shapiro steps, can have a variety of
interesting patterns
such as a devil's staircase related to chaotic dynamics.\cite{Bak1986}

\begin{figure}[h]
\centering
\includegraphics[width=\columnwidth]{fig3_sweep_vyFDC.pdf}
\caption{Average particle velocity  $\langle v_{y} \rangle$
  as a function of $F_{\rm dc}$.
  We let
  $F_{\rm ac}=0.0$ (blue dashed) and 
  $F_{\rm ac}=0.07$ (orange) with $f = 0.01$ 
  as in Fig.~\ref{fig:2_Fd_vy_time}.
  The value of $F_{\rm dc}$ for the first four steps
  %is given,
  correspond with the
  panels (a-d) 
  %indicating a system
  plotted in Fig.~\ref{fig:4_phase}.
  %[DM - clarified?  xx don't understand meaning of indicating a system
  %plotted in Fig.~\ref{fig:4_phase} xx]
}
\label{fig:3_sweep_vyFDC}
\end{figure}

  To study synchronization patterns, 
  it is useful to compare
  mode-locked quantities 
  in a two-dimensional phase plot. 
  For a driven pendulum confined to a single potential well,
  an appropriate
  phase space is the particle velocity $v_y$ versus the position $y$.  
  For a particle driven 
  through multiple identical wells 
  we define phase variables 
  to account for the net increase in the position.
  The phase position is
  \begin{equation}
    \phi(t) = 2\pi [y(t)-\langle v_y \rangle t]/\lambda,
  \end{equation}
  centered about the average particle displacement $\langle v_y \rangle t$
  and normalized by the substrate period $\lambda$.\cite{Juniper2015}
  The phase velocity is
  \begin{equation}
    \dot{\phi}(t) =2\pi [v_y(t)-\langle v_y \rangle] /\lambda.  
  \end{equation}
  For zero landscape force, 
  the phase velocity  is zero when $F_{\rm d}(t) = F_{\rm dc}$.
  %[DM rewrite]
  If the system
  is strictly mode-locked,
  the particle velocity
  recurs at a particular spatial location, and 
  %If the velocity
  %is mode-locked to 
  %a spatial location,
  %[xx don't understand meaning of the velocity
  %is mode-locked to 
  %a spatial location xx]
  a closed loop appears in  
  phase space. 
  A 1:1 mode appears as a circle or oval. 
  Nodes appear 
  for higher modes,
  sometimes forming figure-eights
  or other recognizable patterns.
  A system that is nearly phase locked
  will appear as an almost closed loop.
  Such quasiperiodic systems are
  not fully synchronized
  so the position-velocity relation
  shifts in time as in Fig.~\ref{fig:4_phase}.

  In Fig.~\ref{fig:4_phase}
  we plot $\dot{\phi}(t)$ versus $\phi(t)$
  for increasing 
  $F_{\rm dc}$, 
  with the remaining parameters fixed as in Fig.~\ref{fig:2_Fd_vy_time}.
 %[xx this sentence not helpful, what is plotted is clear: For clarity the plots are normalized by the prefactor $2\pi$.  xx]
  For
  $F_{\rm dc} = 0.04$
  the phase plot   in Fig.~\ref{fig:4_phase}(a) with is an
  % small [xx not clear what small means? can we eliminate this word? xx]
  asymmetric curve.
  A tail appears
  due to the initial transient
  motion of the particle.
  The particle is confined to a single
  substrate minima,
  and has no net velocity.
  The asymmetry is caused by bias induced by $F_{\rm dc}$.
  In Fig.~\ref{fig:4_phase}(b)
  with $F_{\rm dc} = 0.07$
  the phase loop is a symmetric triangular shape,
  indicating a 1:1 match between the
  particle motion and velocity consistent with 
  Fig.~\ref{fig:2_Fd_vy_time}(a).
  As $F_{\rm dc}$ is
  increased,
  nodes form in the phase diagram, which
   occur due to repeated values
  of $\dot{\phi}$ over multiple phase positions.
  In Fig.~\ref{fig:4_phase}(c)
  with $F_{\rm dc} = 0.1$
  two nodes form.
  The particle moves a distance $2\lambda$
  during one time period.
For $F_{\rm dc} = 0.125$ as in Fig.~\ref{fig:4_phase}(d),
  three nodes form as the particle moves across $3\lambda$.

    \begin{figure}[h!]
      \centering
      \includegraphics[width=\columnwidth]{fig4_phase.pdf}
      \caption{
        Phase plot of $\dot{\phi(t)}/(2\pi)$ versus $\phi(t)/(2\pi)$.
        The particle is driven with $F_{\rm dc}$ equal to  (a) $0.04$, (b) $0.07$, (c) $0.1$, and (d) $0.125$.  These values are denoted as black circles
        in Fig.~\ref{fig:3_sweep_vyFDC}. 
      The other parameters
      are $F_{\rm ac} = 0.07$, $f=0.01$, and $A_p = 0.1$
      as in Fig.~\ref{fig:2_Fd_vy_time}.}
      \label{fig:4_phase}
    \end{figure}

@@

\begin{problem}{Write your own MD code}
  \label{ex:euler}
%[xx you didn't really discuss other integration algorithms, you just mentioned some names That's OK. Instead you might want to make this problem 1 and guide  readers to write their own program. Then mention that a program is available, which you only mention in the abstract xx]

  To calculate the position of the particle we
  numerically integrate the equation of motion using
  the standard definition of velocity
  $\vec{v} = d\vec{r}/dt$ 
  using the 
  Euler method. 
  The equation of motion provides
  a direct calculation for the particle velocity,
  as demonstrated in Problem~\ref{ex:n2l}.  
  The Euler method is effective for solving linear ordinary differential  equations
  of the form
  $dy/dt = f(t,y(t))$ with the initial condition $y(t_0) = y_0$.
  The solution is calculated algorithmically
  by stepping in time through $n$ integer steps
  $t_n = t_0 + n \Delta t$.
  At each subsequent step the new
  value for $y$ is calculated as a  solution of a map using
  discrete times 
  $y_{n+1} = y_n + f(t_n,y_n)$.
  Apply the Euler method to %our equation of motion
  Eq.~\ref{eq:motion}
  %[DM - resolved xx refer to specific equation by label xx]
  to determine  the analytic expression
  for the position of a particle
  $y_n$ at the $nth$ time step.

  The Euler algorithm can be applied to
  calculate reasonable numerical solutions to 
  nonlinear
  equations if the time step $\Delta t$
  is kept sufficiently small.\cite{Newman}
  In our simulations we use the time step $\Delta t = 0.1$
  and find no change in the solution
  when we decrease the time step to smaller values.
  In simulations of
  many interacting particles,
  a small
  time step is essential for accurate results.
  Particle-particle interactions are typically nonlinear,
  so that the interparticle force changes significantly over small distances and the simple Euler algorithm is insufficient.

  A working example of this code is available
  in Ref.~\cite{supplemental}

\end{problem}

@@

  %-------------------------------------------------------------------------

@@

\item [(b)] If $m/\gamma$ is small, the
  particle's acceleration can be ignored
  entirely.
Using Newton's second law for
  a small momentum relaxation time, 
  show that a particle confined to a landscape exerting force
  $F_\ell(\vec{r})$ subject to a time dependent drive $F_{\rm d}(t)$
  can be modeled by the equation of motion described in 
  Eq.~\eqref{eq:motion}.
  \end{enumerate}
\end{problem}

%--------------------------------------------------------------------------
